---
title: "Simulating the mother groups DGP "
date: "2025-05-02"
author: "Hans Henrik Sievertsen"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
    toc: true
    toc_float:
      toc_collapsed: true
---


# 1 Goal

The goal is to simulate a data generating process (DGP) mimicking the mother group assignment in Denmark. It is used to understand the role of fixed effects and the role of exclusion bias.

# 2 Preparation

We set seed, clear the workspace and load libraries. 

##### Set seed
```{r}
#| label: set-seed
#| include: true
#| code-fold: false

rm(list=ls())
set.seed(1909)
```

##### Load libriaries
```{r}
#| label: load-libraries
#| include: true
#| code-fold: false
#| warning: false
#| message: false

library(ggplot2)      # To create charts
library(patchwork)    # To mix charts
library(dplyr)        # Data manipulation tools
library(lubridate)    # Date manipulation tools
library(fixest)       # Estimation
library(data.table)   # Data manipulation tools
library(modelsummary) # Create tables 
library(lmtest)       # Hypotheses testing
```




# 3 The Data Generating Process (DGP)

## 3.1 Defining the DGP


We here create a function that simulates a dataset mimicking the DGP we have in mind. The key features we focus on are

1. The role of unobservables (in group assignments and outcomes)

2. The assignment mechanism

3. The peer effects

**Re 1.**: We assume there is an individual unobservable $U$ that is a combination of  a unique individual effect and a district level effect. This unobservable affects variables like labor supply, income, parental labor supply and age. But *not* parity and child dob. 

**Re 2.**: The assignment is only strict based on district. THat is, *within each district*, nurses *sort* the mothers by

1. Child date of birth (days)

2. Mother age (integers)

3. Parity (binary)

We add a bit of noise to the sorting to mimick the real world where there will be errors, mothers coming in during the process etc. The nurse then simply starts filling mother groups by counting to five. That is, we *restrict mother groups to the size of 5*.

**Re 3.**: True peer effects are tricky to simulate. Here we cheat a bit by ignoring the multiplication effect/reflect effect. That is we:

- Define a latent level of mother labor supply depending on grand parent labor supply, the unobservables, etc. 

- Calculate the mother group leave one out average latent labor supply. 

- Calculate the *actual* labour supply by adding the peer effect to the latent effect.


That is it for now. 


```{r}
#| label: define-DGP
#| include: true
#| code-fold: true
#| warning: false

DGP_fast <- function(n = 70000, n_districts = 500) {
    
    # District identifier
    id_district <- sample(1:n_districts, n, replace = TRUE)
    # District unobserved effect
    district_effect <- rnorm(n_districts, mean = 0, sd = 1)
    # Individual unobserved effect
    U <- district_effect[id_district] + rnorm(n)
    # Covariates, note that parity and dob are unrelated to u, age, education, income, gm labour supply are not 
    x_age <- round(18 + 10 * runif(n) + 2 * U + rnorm(n))  # now depends on U
    x_age <- pmax(pmin(x_age, 40), 18)  # keep age between 18 and 40
    x_parity <- sample(1:2, n, replace = TRUE)
    x_schooling <- round(9 + 0.2 * x_age + 0.8 * U + rnorm(n))
    x_income <- round(200000 + 5000 * x_schooling + 10000 * U + rnorm(n, 0, 20000))
    x_gm_lsupply <- pmax(0, 20 + 0.5 * U + rnorm(n, 0, 5))
    x_dob <- sample(seq(as.Date("2010-01-01"), as.Date("2017-12-31"), by = "day"), n, replace = TRUE)
    # Create data frame
    dt <- data.table(id_mother = 1:n, id_district, x_age, x_parity,
                     x_schooling, x_income, x_gm_lsupply, x_dob)
    # Assign mother groups
    setorder(dt, id_district, x_dob, x_age,x_parity)
    dt[, order_original := .I , by = id_district]
    # Add small random noise to row numbers
    dt[, order_noisy := .I + rnorm(.N, mean = 0, sd = 1)]
    # Sort by the noisy key
    setorder(dt, id_district,order_noisy)
    dt[, order_noisy := .I , by = id_district]
    # Assign groups
    dt[, count := 1:.N, by = id_district]
    dt[, group_id := floor((count - 1) / 5)]
    dt[, id_mother_group := paste(id_district, group_id, sep = "_")]
    
    # Create fixed effects
    dt[, x_quarter := lubridate::quarter(x_dob, with_year = TRUE)]
    dt[, x_month := lubridate::month(x_dob)]
    dt[, x_year := lubridate::year(x_dob)]
    dt[, x_young := as.integer(x_age < 30)]
    dt[, x_age5 := floor(x_age/5)*5]
    # Leave-one-out variables
    dt[, `:=`(
      sum_schooling = sum(x_schooling),
      sum_income = sum(x_income),
      sum_gm=sum(x_gm_lsupply),
      n_group = .N
    ), by = id_mother_group]
    dt[, x_schooling_g := (sum_schooling - x_schooling) / (n_group - 1)]
    dt[, x_income_g := (sum_income - x_income) / (n_group - 1)]
    dt[, x_gm_lsupply_g := (sum_gm - x_gm_lsupply) / (n_group - 1)]
    
    # Create labor supply (our main outcome)
    # Initial (baseline) labor supply without peer effect
    dt[, lsupply_base := pmin(40, pmax(0, 
                                       20 +         # base level
                                         0.5 * U +     # unobserved factor
                                         0.3 * x_gm_lsupply +  # grandmother's influence
                                         rnorm(.N, 0, 5)  # noise
    ))]
    # Leave-one-out peer effect within mother group
    dt[, `:=`(
      sum_lsupply = sum(lsupply_base),
      n_group = .N
    ), by = id_mother_group]
    
    dt[, x_lsupply_g := (sum_lsupply - lsupply_base) / (n_group - 1)]
    # Final labor supply including peer effect
    dt[, y_lsupply := pmin(40, pmax(0,
                                    lsupply_base +
                                      0.2 * x_lsupply_g  # peer effect coefficient
    ))]
    
    # Keep only relevant columns
    dt<-dt[, .SD, .SDcols = grep("^id_|^x_|^gm_|^y_", names(dt), value = TRUE)]
      return(dt)
}

```


## 3.2 Showing the DGP


### 3.2.1 The group assignments 

Let us have a look at the data generated by the DGP function defined above. We first focus on the group assignments. The variables are sorted so that without noise they should be perfectly assigned to groups. The group assignment mostly follows this sorting, but with some discrepancies. This seems realistic. 


```{r}
#| label: show data
#| include: true
#| code-fold: true
#| warning: false
#| cache: true


df<-DGP_fast()
setorder(df, id_district, x_parity, x_dob, x_age)
head(df[, .SD, .SDcols = c("id_district","id_district", "x_dob", "x_age","x_parity", "id_mother_group","y_lsupply")],n=20)


```


### 3.2.2 The "first" stage


We can look at many aspects of the data, but as a starter, let us have a look at the link between grand mother labor supply and own labor supply. This looks as expected (but I am puzzled about the NAs, anyway).

```{r}
#| label: show first stage
#| include: true
#| code-fold: true
#| warning: false
#| cache: true


# Collapse data
df_collapsed<-df%>%
          mutate(x_gm_lsupply_round=round(x_gm_lsupply))%>%       # Round grand mother labor supply
                            group_by(x_gm_lsupply_round)%>%       # Group by the rounded labor supply
                            summarize(y_lsupply=mean(y_lsupply,na.rm = T))  # Collapsed 

# Create plot
ggplot(df_collapsed,aes(x=x_gm_lsupply_round,y=y_lsupply))+
      geom_point()+
      theme_minimal()
      

```



### 3.2.3 Summary stats

Finally, let us have a look at some summary stats. There are some variables that could be better defined, but overall it looks fine.


```{r}
#| label: show sum stats
#| include: true
#| code-fold: true
#| warning: false
#| cache: true


datasummary_skim(df)
      

```



# 4 Check for balance

## 4.1 The traditional balance check.

We want to know whether group peers are as-good-as-random assignment. Our key variable is labour supply. So let us check whether the average grand mother labour supply is related to our own mother's labour supply. It should not be (once we adjust for assignment to group variables). We run three regressions

1. Own grandmother labor supply on average peer grand mother labor supply

2. Own grandmother labor supply on average peer grand mother labor supply, conditional on district and mother age fixed effects.

3. Own grandmother labor supply on average peer grand mother labor supply, conditional on district,  mother age, child quarter of birth and parity fixed effects.



```{r}
#| label: balance_reg_approach_reg
#| include: true
#| code-fold: true
#| warning: false
#| cache: true 


m1<-feols(x_gm_lsupply~x_gm_lsupply_g,data=df)
m2<-feols(x_gm_lsupply~x_gm_lsupply_g|id_district+x_age5,data=df)
m3<-feols(x_gm_lsupply~x_gm_lsupply_g|id_district+x_age5+x_quarter+x_parity,data=df)

modelsummary(list(m1,m2,m3),
             coef_omit = "Intercept",
              stars = TRUE)
```

We observe 

- In the raw specification (column (1)), there is a significant positive corellation, as we would expect.

- Once we condition on fixed effects the correlation becomes negative and significant. This is also what we would expect, given the exclusion bias issue. And remember in this case we *know that assignment is only based on the variables we condition on*! 


## 4.2 An alternative approach: check distribution of p-values!

We now test the use of an alternative approach, where we check whether mother group indicators can predict mother groups within the cells where the assignment is done. So basically the idea is that once we only look with a district at similar aged mothers, similar parity and similar child dob, the exact group you get assigned to should be random. Therefore, running a regression of own characteristics on indicators for mother groups, the p-value on the joint F-test for the hypothesis that all indicators could be excluded should be uniformly distributed. 


### 4.2.1 By district only

First we assume that the assignment cell is district We  first loop over cells and districts and run all possible regressions and save the p-values. 

```{r}
#| label: approach2-cell-is-district
#| include: true
#| code-fold: true
#| warning: false
#| cache: true 



# datafarme to save results
df_results<-data.table(id_district=rep(unique(df$id_district),times=4),
                       p=NA,
                       var=rep(c("x_schooling","x_income","x_age", "x_gm_lsupply", "x_parity"),each=length(rep(unique(df$id_district))))
                               )

# Loop over variables
for (v in c("x_schooling","x_income","x_age", "x_gm_lsupply", "x_parity")){
  # Loop over all districts:
  for (i in unique(df$id_district)){
    # select sample
    filtered_df <- df[df$id_district== i, ]
     
    # Model without fixed effects
    mod1 <- feols(as.formula(paste(v, "~ 1")), data = filtered_df)
    
    # Model with fixed effects
    mod2 <- feols(as.formula(paste(v, "~ 1 | id_mother_group")), data = filtered_df)
    
      # Get residual sum of squares
    rss1 <- sum(resid(mod1)^2)
    rss2 <- sum(resid(mod2)^2)
    
    # Get degrees of freedom
    df1 <- df.residual(mod1)
    df2 <- df.residual(mod2)
    
    # Compute F-statistic
    num_df <- df1 - df2  # numerator degrees of freedom (number of added params)
    den_df <- df2        # denominator degrees of freedom
    f_stat <- ((rss1 - rss2) / num_df) / (rss2 / den_df)
        # Compute p-value
    pval <- pf(f_stat, df1 - df2, df2, lower.tail = FALSE)
    # Store results
    df_results$p[df_results$id_district == i & df_results$var == v] <- pval
  }
  print(v)
}

```

As we now have the data, we can look at plots of the p-values


```{r}
#| label: approach2-cell-is-district-vis
#| include: true
#| code-fold: true
#| warning: false
#| cache: true 

# Sort by var and p
setorder(df_results, var, p)

# Compute relative rank within each var group
df_results[, relrank := rank(p, ties.method = "first") / .N, by = var]

# Plot

vis1<-ggplot(df_results, aes(x = relrank, y = p, color = var)) +
   geom_segment(aes(x = 0, xend = 1, y = 0, yend =1),color="black")+
  geom_line() +
  labs(x = "Relative Rank", y = "p-value", color = "Variable") +
  theme_minimal()+
  labs(title="(a) assignment by district")+
  theme(legend.position = "top")

vis1
```

If the p-values are uniformly distributed all lines should be 45 degree lines. They are clearly not.

### 4.2.2 By district and age

We now repeat the exercise by age and district groups

```{r}
#| label: approach2-cell-is-district-and-age
#| include: true
#| code-fold: true
#| warning: false
#| cache: true 


df[, group := .GRP, by = .(id_district,x_age5)]


# datafarme to save results
df_results<-data.table(group=rep(unique(df$group),times=5),
                       p=NA,
                       var=rep(c("x_schooling","x_income","x_age", "x_gm_lsupply","x", "x_parity"),each=length(rep(unique(df$group))))
                               )
# Loop over variables
for (v in c("x_schooling","x_income","x_age", "x_gm_lsupply")){
  # Loop over all districts:
  for (i in unique(df$group)){
    # select sample
    filtered_df <- df[df$group== i, ]
    
    # Skip if the variable is constant or NA
    y <- filtered_df[[v]]
    if (length(unique(filtered_df$id_mother_group)) <10) next
    if (length(unique(y[!is.na(y)])) <= 1) next  # Skip constant or all-NA
    
    
    # Model without fixed effects
    mod1 <- lm(as.formula(paste(v, "~ 1")), data = filtered_df)
    
    # Model with fixed effects
    mod2 <- lm(as.formula(paste(v, "~ factor(id_mother_group)")), data = filtered_df)
  
    # Compare models
   # Get residual sum of squares
    rss1 <- sum(resid(mod1)^2)
    rss2 <- sum(resid(mod2)^2)
    
    # Get degrees of freedom
    df1 <- df.residual(mod1)
    df2 <- df.residual(mod2)
    
    # Compute F-statistic
    num_df <- df1 - df2  # numerator degrees of freedom (number of added params)
    den_df <- df2        # denominator degrees of freedom
    f_stat <- ((rss1 - rss2) / num_df) / (rss2 / den_df)
    
    # Compute p-value
    pval <- pf(f_stat, df1 - df2, df2, lower.tail = FALSE)
    # Store results
    df_results$p[df_results$group == i & df_results$var == v] <- pval
  }
  print(v)
}

```

We now create a plot, like before: 


```{r}
#| label: approach2-cell-is-district-and-age-vis
#| include: true
#| code-fold: true
#| warning: false
#| cache: true 


# Remove rows with NA in p
df_results <- df_results[!is.na(p)]

# Sort by var and p
setorder(df_results, var, p)

# Compute relative rank within each var group
df_results[, relrank := rank(p, ties.method = "first") / .N, by = var]


# Plot
vis2<-ggplot(df_results, aes(x = relrank, y = p, color = var)) +
  geom_line() +
  labs(x = "Relative Rank", y = "p-value", color = "Variable") +
  theme_minimal()+
  xlim(0,1)+ylim(0,1)+
  labs(title="(b) assignment by district & age5")+
  theme(legend.position = "top")

vis2
```

And now it is perfectly blanced. Great! This suggests that it is fairly balanced condition on age and district, but we need to solve the exclusion bias issue.







# 5 Solving the exclusion bias issue

## 5.1 Illustrating that the problem is real

We only showed that in one regression, the coefficient on average peer grandmother labor supply becomes significantly negative, when we condition on district and age5. But this could just be random chance. Naturally, because we generate the data we could also just simulate N DGPs and show that this is not just by chance, but we cannot translate that assessment to the real data.  To illustrate the problem more clearly, we therefore instead randomly reshuffle the mother groups and repeat the regression. In that regression the correlation should be 0 because we reshuffled and we can also do that on the real data. 

```{r}
#| label: reshuffle-mother-groups-and-test
#| include: true
#| code-fold: true
#| warning: false
#| cache: true 


#Iterations
I=10000
# Dataset
df_sim<-data.frame(i=1:I,beta=NA,type="Raw")

for (i in 1:I){
  # random order
  df[, random :=rnorm(.N, mean = 0, sd = 1)]
  # reorder
  setorder(df, id_district, x_age5,random)
  # Assign groups
  df[, count := 1:.N, by = id_district]
  df[, group_id := floor((count - 1) / 5)]
  df[, id_mother_group_new := paste(id_district, group_id, sep = "_")]
  # Recalculate leave one out mean
   df[, `:=`(
      sum_gm=sum(x_gm_lsupply),
      n_group = .N
    ), by = id_mother_group_new]
    df[, x_gm_lsupply_gshuffle := (sum_gm - x_gm_lsupply) / (n_group - 1)]
    
  # Save estimate
    res<-feols(x_gm_lsupply~x_gm_lsupply_gshuffle|id_district+x_age5,data=df,notes=FALSE)
    # Extract coefficient
    beta <- summary(res)$coeftable["x_gm_lsupply_gshuffle", "Estimate"]
    # Store beta-value in results dataset
    df_sim$beta[i] <- beta
    
      # Print
        if (i%%50==0){
             #  cat(paste(i,"\n"))
        }
        else{
             #  cat("*")
        }
    
}

```


Let is show the distribution in a chart. 

```{r}
#| label: balance-plot
#| include: true
#| code-fold: true
#| warning: false


# Store main
main<-feols(x_gm_lsupply~x_gm_lsupply_g|id_district+x_age5,data=df)
beta <- summary(main)$coeftable["x_gm_lsupply_g", "Estimate"]



fig1<-ggplot(df_sim,aes(x=beta))+geom_histogram(color="black",fill="white") +
  theme_minimal()+geom_vline(xintercept = 0)+xlim(-0.1,0.1)+labs(title="(a) raw")

fig1
```

The distribution is clearly not centered around 0, as we would expect!


## 5.2 Approach 1: control for pool level 

The first approach uses the idea of controlling for the pool level. We simply control for leave one out average of grandmother labor supply in the entire district. One could also do it by district and age, but let us start by district. 



```{r}
#| label: simulate-pool-control
#| include: true
#| code-fold: true
#| warning: false
#| cache: true

#Iterations
I=10000
# Dataset
df_sim_pool<-data.frame(i=1:I,beta=NA,type="Raw")

 # Calculate pool level leave one out mean
  df[, group := .GRP, by = .(id_district)]
   df[, `:=`(
      sum_gm=sum(x_gm_lsupply),
      n_group = .N
    ), by = group]
  df[, x_gm_lsupply_da := (sum_gm - x_gm_lsupply) / (n_group - 1)]
    

  
for (i in 1:I){
  # random order
  df[, random :=rnorm(.N, mean = 0, sd = 1)]
  # reorder
  setorder(df, id_district, x_age5,random)
  # Assign groups
  df[, count := 1:.N, by = id_district]
  df[, group_id := floor((count - 1) / 5)]
  df[, id_mother_group_new := paste(id_district, group_id, sep = "_")]
  # Recalculate leave one out mean
   df[, `:=`(
      sum_gm=sum(x_gm_lsupply),
      n_group = .N
    ), by = id_mother_group_new]
    df[, x_gm_lsupply_gshuffle := (sum_gm - x_gm_lsupply) / (n_group - 1)]
    
  # Save estimate
    res<-feols(x_gm_lsupply~x_gm_lsupply_gshuffle+x_gm_lsupply_da|id_district+x_age5,data=df,notes=FALSE)
    # Extract coefficient
    beta <- summary(res)$coeftable["x_gm_lsupply_gshuffle", "Estimate"]
    # Store beta-value in results dataset
    df_sim_pool$beta[i] <- beta
    
      # Print
        if (i%%50==0){
           #    cat(paste(i,"\n"))
        }
        else{
           #    cat("*")
        }
    
}

```


Let us

```{r}
#| label: balance-plot-pool
#| include: true
#| code-fold: true
#| warning: false





fig2<-ggplot(df_sim_pool,aes(x=beta))+geom_histogram(color="black",fill="white") +
  theme_minimal()+geom_vline(xintercept = 0)+labs(title="(b) control for pool level")+
  xlim(-0.025,0.025)

fig1+fig2
```

This clearly centers the distribution around 0! Great. That solution works. 





## 5.3 Approach 2: recenter variable based on adding variation in date of birth

Let us know try the recentering approach. We could simply recenter by subtracting the pool level leave one out from the mother group leave one out. That should result in the same as above. However, we will know be more explicit about the pool level and instead simulate pools based on the mother groups you could be assigned to, by adding a bit of noise to the date of birth. 



```{r}
#| label: recenter-sim-14
#| include: true
#| code-fold: true
#| warning: false
#| cache: true

  # Dataset for results
  df_sim_recent<-data.frame(i=1:I,beta=NA,type="Raw")
  
          
 I=1000

  for (i in 1:I){

              

    
  # calculate random 
  J=20
    df_test<-df
  # Empty data table to store results
  df_permute<-data.table(id_mother=rep(df_test$id_mother,times=I),
                        iteration=rep(1:I,each=length(df_test$id_mother),
                        beta=NA)
                  )

   setorder(df_test, id_mother)
  for (j in 1:J){
    # New data of birth
      df_test[, x_dob_new := x_dob + round(runif(.N, min = -14, max = 14))]
    # Assign mother groups
      setorder(df_test, id_district, x_dob_new, x_age,x_parity)
    # Assign groups
      df_test[, count := 1:.N, by = id_district]
      df_test[, group_id := floor((count - 1) / 5)]
      df_test[, id_mother_group_new := paste(id_district, group_id, sep = "_")]
     # Recalculate leave one out mean
     df_test[, `:=`(
        sum_gm=sum(x_gm_lsupply),
        n_group = .N
      ), by = id_mother_group_new]
      df_test[, x_gm_lsupply_gs := (sum_gm - x_gm_lsupply) / (n_group - 1)]
     # Save
      setorder(df, id_mother)
      df_permute$beta[df_permute$iteration == j ] <- df_test$x_gm_lsupply_gs
      
  }
  
  df_collapsed <- df_permute[, .(x_gm_lsupply_gsim14 = mean(beta, na.rm = TRUE)), by = id_mother]


  # now merge on

  df<-merge(df,df_collapsed,by="id_mother")
  df[,x_gm_lsupply_grc :=x_gm_lsupply_g-x_gm_lsupply_gsim14 ]
  
  # Save estimate
    res<-feols(x_gm_lsupply~x_gm_lsupply_grc|id_district+x_age5,data=df,notes=FALSE)
    # Extract coefficient
    beta <- summary(res)$coeftable["x_gm_lsupply_grc", "Estimate"]
    # Store beta-value in results dataset
    df_sim_recent$beta[i] <- beta
    
     # Print
        if (i%%50==0){
         #      cat(paste(i,"\n"))
        }
        else{
      #         cat("*")
        }
      df<-df%>%select(-x_gm_lsupply_gsim14)
}


```



Let us again create a histogram and compare it to the other histograms


```{r}
#| label: balance-plot-recenter-14
#| include: true
#| code-fold: true
#| warning: false





fig3<-ggplot(df_sim_recent,aes(x=beta))+geom_histogram(color="black",fill="white") +
  theme_minimal()+geom_vline(xintercept = 0)+labs(title="(c) recenter w.  dob 14d")+
  xlim(-0.025,0.025)



fig1+fig2+fig3
```

This is not as narrow as (b), and worse, it is not centered. Can we do better?


## 5.4 Approach 3: a more generous recentering based on adding variation in date of birth

Let us know try the recentering approach. We could simply recenter by subtracting the pool level leave one out from the mother group leave one out. That should result in the same as above. However, we will know be more explicit about the pool level and instead simulate pools based on the mother groups you could be assigned to, by adding a bit of noise to the date of birth. 



```{r}
#| label: recenter-sim-90
#| include: true
#| code-fold: true
#| warning: false
#| cache: true

  # Dataset for results
  df_sim_recent<-data.frame(i=1:I,beta=NA,type="Raw")
  
          
 I=1000

  for (i in 1:I){

              

    
  # calculate random 
  J=20
    df_test<-df
  # Empty data table to store results
  df_permute<-data.table(id_mother=rep(df_test$id_mother,times=I),
                        iteration=rep(1:I,each=length(df_test$id_mother),
                        beta=NA)
                  )

   setorder(df_test, id_mother)
  for (j in 1:J){
    # New data of birth
      df_test[, x_dob_new := x_dob + round(runif(.N, min = -90, max = 90))]
    # Assign mother groups
      setorder(df_test, id_district, x_dob_new, x_age,x_parity)
    # Assign groups
      df_test[, count := 1:.N, by = id_district]
      df_test[, group_id := floor((count - 1) / 5)]
      df_test[, id_mother_group_new := paste(id_district, group_id, sep = "_")]
     # Recalculate leave one out mean
     df_test[, `:=`(
        sum_gm=sum(x_gm_lsupply),
        n_group = .N
      ), by = id_mother_group_new]
      df_test[, x_gm_lsupply_gs := (sum_gm - x_gm_lsupply) / (n_group - 1)]
     # Save
      setorder(df, id_mother)
      df_permute$beta[df_permute$iteration == j ] <- df_test$x_gm_lsupply_gs
      
  }
  
  df_collapsed <- df_permute[, .(x_gm_lsupply_gsim90 = mean(beta, na.rm = TRUE)), by = id_mother]


  # now merge on

  df<-merge(df,df_collapsed,by="id_mother")
  df[,x_gm_lsupply_grc :=x_gm_lsupply_g-x_gm_lsupply_gsim90 ]
  
  # Save estimate
    res<-feols(x_gm_lsupply~x_gm_lsupply_grc|id_district+x_age5,data=df,notes=FALSE)
    # Extract coefficient
    beta <- summary(res)$coeftable["x_gm_lsupply_grc", "Estimate"]
    # Store beta-value in results dataset
    df_sim_recent$beta[i] <- beta
    
     # Print
        if (i%%50==0){
         #      cat(paste(i,"\n"))
        }
        else{
      #         cat("*")
        }
      df<-df%>%select(-x_gm_lsupply_gsim90)
}


```

Let us again create a histogram and compare it to the other histograms

```{r}
#| label: balance-plot-recenter-90
#| include: true
#| code-fold: true
#| warning: false





fig4<-ggplot(df_sim_recent,aes(x=beta))+geom_histogram(color="black",fill="white") +
  theme_minimal()+geom_vline(xintercept = 0)+labs(title="(d) recenter w.  dob 90d")+
  xlim(-0.025,0.025)



(fig1+fig2)/(fig3+fig4)
```

This is not as narrow as (b), and worse, it is not centered. Can we do better?
