---
title: "Simulating the mother groups DGP "
date: "2025-03-27"
author: "Hans Henrik Sievertsen"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
---


# 1 Goal

The goal is to simulate a data generating process mimicking the mother group assignment in Denmark. It is used to understand the role of fixed effects and the role of exclusion bias. Focus in on balancing tests and confounding factors. 

# 2 Preparation

We set seed and load libraries. 

##### Set seed
```{r}
#| label: set-seed
#| include: true
#| code-fold: false

set.seed(1909)
```

##### Load libriaries
```{r}
#| label: load-libraries
#| include: true
#| code-fold: false
#| warning: false

library(ggplot2)     # To create charts
library(dplyr)       # Data manipulation tools
library(lubridate)   # Date manipulation tools
library(fixest)      # Estimation
library(data.table)  # Data manipulation tools


```




# 2 Define Data Generating Process

Note that the assignment is only strict based on district, but not based on mother age, parity and date of birth. This is true to the real DGP (I believe), but makes the FE tricky. I.e. a nurse will only assign mothers to mother groups within her own district, but in terms of age, parity and date of birth (of the child) she is not strict, but instead just allocated to the closest neighbour.  

```{r}
#| label: define-DGP
#| include: true
#| code-fold: true
#| warning: false

DGP_fast <- function(n = 70000, n_districts = 500) {
  # District identifier
  id_district <- sample(1:n_districts, n, replace = TRUE)
  # District unobserved effect
  district_effect <- rnorm(n_districts, mean = 0, sd = 1)
  # Individual unobserved effect
  U <- district_effect[id_district] + rnorm(n)
  # Covariates
  x_age <- round(runif(n, 18, 40))
  x_parity <- sample(1:2, n, replace = TRUE)
  x_schooling <- round(9 + 0.2 * x_age + 0.8 * U + rnorm(n))
  x_income <- round(200000 + 5000 * x_schooling + 10000 * U + rnorm(n, 0, 20000))
  gm_lsupply <- pmax(0, 20 + 0.5 * U + rnorm(n, 0, 5))
  x_dob <- sample(seq(as.Date("2010-01-01"), as.Date("2017-12-31"), by = "day"), n, replace = TRUE)
  # Create data frame
  dt <- data.table(id_mother = 1:n, id_district, x_age, x_parity,
                   x_schooling, x_income, gm_lsupply, x_dob)
  # Assign mother groups
  setorder(dt, id_district, x_parity, x_dob, x_age)
  dt[, count := 1:.N, by = id_district]
  dt[, group_id := floor((count - 1) / 5)]
  dt[, id_mother_group := paste(id_district, group_id, sep = "_")]
  # Create fixed effects
  dt[, x_quarter := lubridate::quarter(x_dob, with_year = TRUE)]
  dt[, x_month := lubridate::month(x_dob)]
  dt[, x_year := lubridate::year(x_dob)]
  dt[, x_young := as.integer(x_age < 30)]
  # Leave-one-out calculations
  dt[, `:=`(
    sum_schooling = sum(x_schooling),
    sum_income = sum(x_income),
    n_group = .N
  ), by = id_mother_group]
  dt[, x_schooling_m1 := (sum_schooling - x_schooling) / (n_group - 1)]
  dt[, x_income_m1 := (sum_income - x_income) / (n_group - 1)]
  # Leave-one-out calculations, district, year and month level
  dt[, `:=`(
    sum_schooling = sum(x_schooling),
    sum_income = sum(x_income),
    n_group = .N
  ), by = id_district]
  dt[, x_schooling_d1 := (sum_schooling - x_schooling) / (n_group - 1)]
  dt[, x_income_d1 := (sum_income - x_income) / (n_group - 1)]
  # Recentered IV 
  dt[, x_schooling_m1_recentered := x_schooling_m1-x_schooling_d1]
  # Keep only relevant columns
  dt[, .SD, .SDcols = patterns("^id_", "^x_", "^gm_")]
  return(dt)
}

```


#### Show data
```{r}
#| label: show data
#| include: true
#| code-fold: true
#| warning: false

df<-DGP_fast()
setorder(df, id_district, x_parity, x_dob, x_age)
head(df[, .SD, .SDcols = c("id_district","x_dob","x_age","x_parity","id_mother_group")],n=20)

```

# 3 The conventional balancing check!

So now let us see what happens when we test for balancing.

#### Preparation

First we create a dataset to store results

```{r}
#| label: prepare-iterations
#| include: true
#| code-fold: true
#| warning: false

#Iterations
I=1000
# Dataset
df_results<-data.frame(i=1:I,p=NA,type="1 Conventional")
df_results_pool<-data.frame(i=1:I,p=NA,type="2 Control for pool level")
df_results_controls<-data.frame(i=1:I,p=NA,type="3 Flexible controls instead of FE")
df_results_controls_pool<-data.frame(i=1:I,p=NA,type="4 Pool level & flexible controls")
df_results_recenter<-data.frame(i=1:I,p=NA,type="5 Recentered")
df_results_placebo<-data.frame(i=1:I,p=NA,type="6 Placebo")
```

#### Simulate

Now we test for balance using 3 different approach

1. We regress own schooling on peer schooling: $schooling_i=\alpha_0+\alpha_1 schooling_{i-1}+DistrictFE_i+YoungAgeFE_i+ParityFE_i+MonthOfBirthFE_i+u_i$.

2. We regress own schooling on peer schooling, including pool leave one out level schooling:
$schooling_i=\alpha_0+\alpha_1 schooling_{i-1}+\alpha_2 DistrictSchooling_{i-1}+DistrictFE_i+YoungAgeFE_i+ParityFE_i+MonthOfBirthFE_i+u_i$.

3. Like 2, but with FE for year and month and controls for mother age:
$schooling_i=\alpha_0+\alpha_1 schooling_{i-1}+\alpha_2 DistrictSchooling_{i-1}+\alpha_3 AgeFE_i+DistrictFE_i+ParityFE_i+YearFE_i+MonthFE_i+u_i$.

And we then compare that to a placebo regression (where we added a random variable, ideally we would randomly reallocate variables, but one step at the time)


```{r}
#| label: simulate-balancing
#| include: true
#| code-fold: true
#| warning: false
#| cache: true

for (i in 1:I){
    # Generate DGP
    mydata<-DGP_fast()
    # Add random variable
    mydata$x=rnorm(n=nrow(mydata),0,1)
    # 1 Conventional
        res<-feols(x_schooling~x_schooling_m1|id_district+x_quarter+x_young+x_parity,data=mydata,notes=FALSE)
        # Extract p-value from F-test
        p_value <- summary(res)$coeftable["x_schooling_m1", "Pr(>|t|)"]
        # Store p-value in results dataset
        df_results$p[i] <- p_value
    # 2 Control for pool level
        res<-feols(x_schooling~x_schooling_m1+x_schooling_d1|id_district+x_quarter+x_young+x_parity,data=mydata,notes=FALSE)
        # Extract p-value from F-test
        p_value <- summary(res)$coeftable["x_schooling_m1", "Pr(>|t|)"]
        # Store p-value in results dataset
        df_results_pool$p[i] <- p_value
     #  3 Control for pool level, flexible controls 
        res<-feols(x_schooling~x_schooling_m1+x_age+x_parity|id_district+x_year+x_month,data=mydata,notes=FALSE)
        # Extract p-value from F-test
        p_value <- summary(res)$coeftable["x_schooling_m1", "Pr(>|t|)"]
        # Store p-value in results dataset
        df_results_controls$p[i] <- p_value
    #  4 Control for pool level, flexible controls
        res<-feols(x_schooling~x_schooling_m1+x_schooling_d1+x_age+x_parity|id_district+x_year+x_month,data=mydata,notes=FALSE)
        # Extract p-value from F-test
        p_value <- summary(res)$coeftable["x_schooling_m1", "Pr(>|t|)"]
        # Store p-value in results dataset
        df_results_controls_pool$p[i] <- p_value
    #  5 Recentered IV
        res<-feols(x_schooling~x_schooling_m1_recentered|id_district+x_quarter+x_young+x_parity,data=mydata,notes=FALSE)
        # Extract p-value from F-test
        p_value <- summary(res)$coeftable["x_schooling_m1_recentered", "Pr(>|t|)"]
        # Store p-value in results dataset
        df_results_recenter$p[i] <- p_value
    #  6 Placebo
        res<-feols(x_schooling~x|id_district+x_quarter+x_young+x_parity,data=mydata,notes=FALSE)
        # Extract p-value from F-test
        p_value <- summary(res)$coeftable["x", "Pr(>|t|)"]
        # Store p-value in results dataset
        df_results_placebo$p[i] <- p_value
    
    # Print
        if (i%%50==0){
               cat(paste(i,"\n"))
        }
        else{
               cat("*")
        }
    
}
```

#### Charts 

Now we create charts of p values. They should be uniformly distributed. We start with a  histogram.

```{r}
#| label: chart-histogram
#| include: true
#| code-fold: true
#| warning: false

df<-rbind(df_results,df_results_placebo,df_results_controls,df_results_pool,df_results_controls_pool,df_results_recenter)

ggplot(df,aes(x=p,group=type,fill=type))+
    geom_histogram(alpha=0.5,color="black")+
  scale_fill_viridis_d() +
  facet_wrap(~type)+
  theme_minimal()+labs(color=" ",fill=" ")+
  theme(legend.position = "none")
```


It might be easier to read in a plot of CDFS 

```{r}
#| label: chart-cdf
#| include: true
#| code-fold: true
#| warning: false
#| 
df<-rbind(df_results,df_results_placebo,df_results_controls,df_results_pool,df_results_controls_pool,df_results_recenter)%>%
  arrange(type,p)%>%
  group_by(type)%>%mutate(rank=row_number()/n())

ggplot(df,aes(x=p,group=type,y=rank,color=type))+
    geom_point(alpha=0.2)+
  scale_color_viridis_d() +
  theme_minimal()+labs(color=" ",fill=" ")+
  theme(legend.position = "top")+ 
guides(color = guide_legend(nrow = 2))
```


