---
title: "Simulating the mother groups DGP "
date: "2025-03-27"
author: "Hans Henrik Sievertsen"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
---


# 1 Goal

The goal is to simulate a data generating process mimicking the mother group assignment in Denmark. It is used to understand the role of fixed effects and the role of exclusion bias. Focus in on balancing tests and confounding factors. 

# 2 Preparation

We set seed and load libraries. 

##### Set seed
```{r}
#| label: set-seed
#| include: true
#| code-fold: false

set.seed(1909)
```

##### Load libriaries
```{r}
#| label: load-libraries
#| include: true
#| code-fold: false
#| warning: false

library(ggplot2)     # To create charts
library(dplyr)       # Data manipulation tools
library(lubridate)   # Date manipulation tools
library(fixest)      # Estimation
library(data.table)  # Data manipulation tools
library(modelsummary)
library(lmtest)
```




# 2 Define Data Generating Process

Note that the assignment is only strict based on district, but not based on mother age, parity and date of birth. This is true to the real DGP (I believe), but makes the FE tricky. I.e. a nurse will only assign mothers to mother groups within her own district, but in terms of age, parity and date of birth (of the child) she is not strict, but instead just allocated to the closest neighbour.  

```{r}
#| label: define-DGP
#| include: true
#| code-fold: true
#| warning: false

DGP_fast <- function(n = 70000, n_districts = 500) {
    
    # District identifier
    id_district <- sample(1:n_districts, n, replace = TRUE)
    # District unobserved effect
    district_effect <- rnorm(n_districts, mean = 0, sd = 1)
    # Individual unobserved effect
    U <- district_effect[id_district] + rnorm(n)
    # Covariates, note that parity and dob are unrelated to u, age, education, income, gm labour supply are not 
    x_age <- round(18 + 10 * runif(n) + 2 * U + rnorm(n))  # now depends on U
    x_age <- pmax(pmin(x_age, 40), 18)  # keep age between 18 and 40
    x_parity <- sample(1:2, n, replace = TRUE)
    x_schooling <- round(9 + 0.2 * x_age + 0.8 * U + rnorm(n))
    x_income <- round(200000 + 5000 * x_schooling + 10000 * U + rnorm(n, 0, 20000))
    x_gm_lsupply <- pmax(0, 20 + 0.5 * U + rnorm(n, 0, 5))
    x_dob <- sample(seq(as.Date("2010-01-01"), as.Date("2017-12-31"), by = "day"), n, replace = TRUE)
    # Create data frame
    dt <- data.table(id_mother = 1:n, id_district, x_age, x_parity,
                     x_schooling, x_income, x_gm_lsupply, x_dob)
    # Assign mother groups
    setorder(dt, id_district, x_dob, x_age,x_parity)
    dt[, order_original := .I , by = id_district]
    # Add small random noise to row numbers
    dt[, order_noisy := .I + rnorm(.N, mean = 0, sd = 1)]
    # Sort by the noisy key
    setorder(dt, id_district,order_noisy)
    dt[, order_noisy := .I , by = id_district]
    # Assign groups
    dt[, count := 1:.N, by = id_district]
    dt[, group_id := floor((count - 1) / 5)]
    dt[, id_mother_group := paste(id_district, group_id, sep = "_")]
    
    # Create fixed effects
    dt[, x_quarter := lubridate::quarter(x_dob, with_year = TRUE)]
    dt[, x_month := lubridate::month(x_dob)]
    dt[, x_year := lubridate::year(x_dob)]
    dt[, x_young := as.integer(x_age < 30)]
    dt[, x_age5 := floor(x_age/5)*5]
    # Leave-one-out variables
    dt[, `:=`(
      sum_schooling = sum(x_schooling),
      sum_income = sum(x_income),
      sum_gm=sum(x_gm_lsupply),
      n_group = .N
    ), by = id_mother_group]
    dt[, x_schooling_g := (sum_schooling - x_schooling) / (n_group - 1)]
    dt[, x_income_g := (sum_income - x_income) / (n_group - 1)]
    dt[, x_gm_lsupply_g := (sum_gm - x_gm_lsupply) / (n_group - 1)]
    
    # Create labor supply (our main outcome)
    # Initial (baseline) labor supply without peer effect
    dt[, lsupply_base := pmin(40, pmax(0, 
                                       20 +         # base level
                                         0.5 * U +     # unobserved factor
                                         0.3 * x_gm_lsupply +  # grandmother's influence
                                         rnorm(.N, 0, 5)  # noise
    ))]
    # Leave-one-out peer effect within mother group
    dt[, `:=`(
      sum_lsupply = sum(lsupply_base),
      n_group = .N
    ), by = id_mother_group]
    
    dt[, x_lsupply_g := (sum_lsupply - lsupply_base) / (n_group - 1)]
    # Final labor supply including peer effect
    dt[, y_lsupply := pmin(40, pmax(0,
                                    lsupply_base +
                                      0.2 * x_lsupply_g  # peer effect coefficient
    ))]
    
    # Keep only relevant columns
    dt<-dt[, .SD, .SDcols = grep("^id_|^x_|^gm_|^y_", names(dt), value = TRUE)]
      return(dt)
}

```


#### Show data
```{r}
#| label: show data
#| include: true
#| code-fold: true
#| warning: false

df<-DGP_fast()
setorder(df, id_district, x_parity, x_dob, x_age)
head(df[, .SD, .SDcols = c("id_district","id_mother_group","y_lsupply")],n=11)


```

# 3 Check that this looks like our real data

## 3.1 Balance checks

### 3.1.1 Approach 1

```{r}
#| label: balance-one regressions
#| include: true
#| code-fold: true
#| warning: false

summary(feols(x_gm_lsupply~x_gm_lsupply_g|id_district+x_age5,data=df))

```


Okay this looks similar 

Let us try to randomize groups and recalculate and check distribution

```{r}
#| label: balance-one regressions
#| include: true
#| code-fold: true
#| warning: false

#Iterations
I=10000
# Dataset
df_sim<-data.frame(i=1:I,beta=NA,type="Raw")

for (i in 1:I){
  # random order
  df[, random :=rnorm(.N, mean = 0, sd = 1)]
  # reorder
  setorder(df, id_district, x_age5,random)
  # Assign groups
  df[, count := 1:.N, by = id_district]
  df[, group_id := floor((count - 1) / 5)]
  df[, id_mother_group_new := paste(id_district, group_id, sep = "_")]
  # Recalculate leave one out mean
   df[, `:=`(
      sum_gm=sum(x_gm_lsupply),
      n_group = .N
    ), by = id_mother_group_new]
    df[, x_gm_lsupply_gshuffle := (sum_gm - x_gm_lsupply) / (n_group - 1)]
    
  # Save estimate
    res<-feols(x_gm_lsupply~x_gm_lsupply_gshuffle|id_district+x_age5,data=df,notes=FALSE)
    # Extract coefficient
    beta <- summary(res)$coeftable["x_gm_lsupply_gshuffle", "Estimate"]
    # Store beta-value in results dataset
    df_sim$beta[i] <- beta
    
      # Print
        if (i%%50==0){
               cat(paste(i,"\n"))
        }
        else{
               cat("*")
        }
    
}

```


Add actual to chart


```{r}
#| label: balance-plot
#| include: true
#| code-fold: true
#| warning: false


# Store main
main<-feols(x_gm_lsupply~x_gm_lsupply_g|id_district+x_age5,data=df)
beta <- summary(main)$coeftable["x_gm_lsupply_g", "Estimate"]



ggplot(df_sim,aes(x=beta))+geom_histogram(color="black",fill="white") +
  theme_minimal()+geom_vline(xintercept = beta)
```

### 3.1.2 Approach 2


```{r}
#| label: reg on cells


# datafarme to save results
df_results<-data.table(id_district=rep(unique(df$id_district),times=4),
                       p=NA,
                       var=rep(c("x_schooling","x_income","x_age", "x_gm_lsupply", "x_parity"),each=length(rep(unique(df$id_district))))
                               )
# Loop over variables
for (v in c("x_schooling","x_income","x_age", "x_gm_lsupply", "x_parity")){
  # Loop over all districts:
  for (i in unique(df$id_district)){
    # select sample
    filtered_df <- df[df$id_district== i, ]
     
    # Model without fixed effects
    mod1 <- feols(as.formula(paste(v, "~ 1")), data = filtered_df)
    
    # Model with fixed effects
    mod2 <- feols(as.formula(paste(v, "~ 1 | id_mother_group_new")), data = filtered_df)
    
    # Compare models
    ll1 <- logLik(mod1)
    ll2 <- logLik(mod2)
    lr_stat <- 2 * (as.numeric(ll2) - as.numeric(ll1))
    df_fe <- length(unique(filtered_df$id_mother_group_new)) - 1
    pval <- pchisq(lr_stat, df = df_fe, lower.tail = FALSE)
    # Store results
    df_results$p[df_results$id_district == i & df_results$var == v] <- pval
  }
  print(v)
}

```

Now create plot

```{r}

# Sort by var and p
setorder(df_results, var, p)

# Compute relative rank within each var group
df_results[, relrank := rank(p, ties.method = "first") / .N, by = var]


# Plot
library(ggplot2)
ggplot(df_results, aes(x = relrank, y = p, color = var)) +
  geom_line() +
  labs(x = "Relative Rank", y = "p-value", color = "Variable") +
  theme_minimal()
```

Now within age and district cells


```{r}
df[, group := .GRP, by = .(id_district,x_age5)]

# datafarme to save results
df_results<-data.table(group=rep(unique(df$group),times=4),
                       p=NA,
                       var=rep(c("x_schooling","x_income","x_age", "x_gm_lsupply", "x_parity"),each=length(rep(unique(df$group))))
                               )
# Loop over variables
for (v in c("x_schooling","x_income","x_age", "x_gm_lsupply", "x_parity")){
  # Loop over all districts:
  for (i in unique(df$group)){
    # select sample
    filtered_df <- df[df$group== i, ]
    
    # Skip if the variable is constant or NA
    y <- filtered_df[[v]]
    if (length(unique(y[!is.na(y)])) <= 1) next  # Skip constant or all-NA
    
    
    # Model without fixed effects
    mod1 <- feols(as.formula(paste(v, "~ 1")), data = filtered_df)
    
    # Model with fixed effects
    mod2 <- feols(as.formula(paste(v, "~ 1 | id_mother_group_new")), data = filtered_df)
    
    # Compare models
    ll1 <- logLik(mod1)
    ll2 <- logLik(mod2)
    lr_stat <- 2 * (as.numeric(ll2) - as.numeric(ll1))
    df_fe <- length(unique(filtered_df$id_mother_group_new)) - 1
    pval <- pchisq(lr_stat, df = df_fe, lower.tail = FALSE)
    # Store results
    df_results$p[df_results$group == i & df_results$var == v] <- pval
  }
  print(v)
}

```

Now create plot

```{r}

# Sort by var and p
setorder(df_results, var, p)

# Compute relative rank within each var group
df_results[, relrank := rank(p, ties.method = "first") / .N, by = var]


# Plot
library(ggplot2)
ggplot(df_results, aes(x = relrank, y = p, color = var)) +
  geom_line() +
  labs(x = "Relative Rank", y = "p-value", color = "Variable") +
  theme_minimal()
```

# 3 The conventional balancing check

So now let us see what happens when we test for balancing.

#### Preparation

First we create a dataset to store results

```{r}
#| label: prepare-iterations
#| include: true
#| code-fold: true
#| warning: false

#Iterations
I=1000
# Dataset
df_results<-data.frame(i=1:I,p=NA,type="1 Conventional")
df_results_pool<-data.frame(i=1:I,p=NA,type="2 Control for pool level")
df_results_controls<-data.frame(i=1:I,p=NA,type="3 Flexible controls instead of FE")
df_results_controls_pool<-data.frame(i=1:I,p=NA,type="4 Pool level & flexible controls")
df_results_recenter<-data.frame(i=1:I,p=NA,type="5 Recentered")
df_results_placebo<-data.frame(i=1:I,p=NA,type="6 Placebo")
```

#### Simulate

Now we test for balance using 3 different approach

1. We regress own schooling on peer schooling: $schooling_i=\alpha_0+\alpha_1 schooling_{i-1}+DistrictFE_i+YoungAgeFE_i+ParityFE_i+MonthOfBirthFE_i+u_i$.

2. We regress own schooling on peer schooling, including pool leave one out level schooling:
$schooling_i=\alpha_0+\alpha_1 schooling_{i-1}+\alpha_2 DistrictSchooling_{i-1}+DistrictFE_i+YoungAgeFE_i+ParityFE_i+MonthOfBirthFE_i+u_i$.

3. Like 2, but with FE for year and month and controls for mother age:
$schooling_i=\alpha_0+\alpha_1 schooling_{i-1}+\alpha_2 DistrictSchooling_{i-1}+\alpha_3 AgeFE_i+DistrictFE_i+ParityFE_i+YearFE_i+MonthFE_i+u_i$.

And we then compare that to a placebo regression (where we added a random variable, ideally we would randomly reallocate variables, but one step at the time)


```{r}
#| label: simulate-balancing
#| include: true
#| code-fold: true
#| warning: false
#| cache: true  

for (i in 1:I){
    # Generate DGP
    mydata<-DGP_fast()
    # Add random variable
    mydata$x=rnorm(n=nrow(mydata),0,1)
    # 1 Conventional
        res<-feols(x_schooling~x_schooling_g|id_district+x_quarter+x_young+x_parity,data=mydata,notes=FALSE)
        # Extract p-value from F-test
        p_value <- summary(res)$coeftable["x_schooling_g", "Pr(>|t|)"]
        # Store p-value in results dataset
        df_results$p[i] <- p_value
    # 2 Control for pool level
        res<-feols(x_schooling~x_schooling_g+x_schooling_d|id_district+x_quarter+x_young+x_parity,data=mydata,notes=FALSE)
        # Extract p-value from F-test
        p_value <- summary(res)$coeftable["x_schooling_g", "Pr(>|t|)"]
        # Store p-value in results dataset
        df_results_pool$p[i] <- p_value
     #  3 Flexible controls 
        res<-feols(x_schooling~x_schooling_g+x_age+x_parity|id_district+x_year+x_month,data=mydata,notes=FALSE)
        # Extract p-value from F-test
        p_value <- summary(res)$coeftable["x_schooling_g", "Pr(>|t|)"]
        # Store p-value in results dataset
        df_results_controls$p[i] <- p_value
    #  4 Control for pool level, flexible controls
        res<-feols(x_schooling~x_schooling_g+x_schooling_d+x_age+x_parity|id_district+x_year+x_month,data=mydata,notes=FALSE)
        # Extract p-value from F-test
        p_value <- summary(res)$coeftable["x_schooling_g", "Pr(>|t|)"]
        # Store p-value in results dataset
        df_results_controls_pool$p[i] <- p_value
    #  5 Recentered IV
        mydata$x_schooling_g_recentered<-mydata$x_schooling_g-mydata$x_schooling_d
        res<-feols(x_schooling~x_schooling_g_recentered|id_district+x_quarter+x_young+x_parity,data=mydata,notes=FALSE)
        # Extract p-value from F-test
        p_value <- summary(res)$coeftable["x_schooling_g_recentered", "Pr(>|t|)"]
        # Store p-value in results dataset
        df_results_recenter$p[i] <- p_value
    #  6 Placebo
        res<-feols(x_schooling~x|id_district+x_quarter+x_young+x_parity,data=mydata,notes=FALSE)
        # Extract p-value from F-test
        p_value <- summary(res)$coeftable["x", "Pr(>|t|)"]
        # Store p-value in results dataset
        df_results_placebo$p[i] <- p_value
    
    # Print
        if (i%%50==0){
               cat(paste(i,"\n"))
        }
        else{
               cat("*")
        }
    
}
```

#### Charts 

Now we create charts of p values. They should be uniformly distributed. We start with a  histogram.

```{r}
#| label: chart-histogram
#| include: true
#| code-fold: true
#| warning: false
#| cache: true

df<-rbind(df_results,df_results_placebo,df_results_controls,df_results_pool,df_results_controls_pool,df_results_recenter)

ggplot(df,aes(x=p,group=type,fill=type))+
    geom_histogram(alpha=0.5,color="black")+
  scale_fill_viridis_d() +
  facet_wrap(~type)+
  theme_minimal()+labs(color=" ",fill=" ")+
  theme(legend.position = "none")
```


It might be easier to read in a plot of CDFS 

```{r}
#| label: chart-cdf
#| include: true
#| code-fold: true
#| warning: false

df<-rbind(df_results,df_results_placebo,df_results_controls,df_results_pool,df_results_controls_pool,df_results_recenter)%>%
  arrange(type,p)%>%
  group_by(type)%>%mutate(rank=row_number()/n())

ggplot(df,aes(x=p,group=type,y=rank,color=type))+
    geom_point(alpha=0.2)+
  scale_color_viridis_d() +
  theme_minimal()+labs(color=" ",fill=" ")+
  theme(legend.position = "top")+ 
guides(color = guide_legend(nrow = 2))
```



# 4 The outcome

```{r}
#| label: prepare-iterations-outc
#| include: true
#| code-fold: true
#| warning: false

#Iterations
I=1000
# Dataset
df_results<-data.frame(i=1:I,bias=NA,type="1 Conventional")
df_results_recenter<-data.frame(i=1:I,bias=NA,type="2 Recentered")

```


```{r}
#| label: outcome-sim
#| include: true
#| code-fold: true
#| warning: false
#| cache: true

for (i in 1:I){
      # Generate DGP
    mydata<-DGP_fast()
    # 1 conventional reg
       res<-feols(y_lsupply~x_lsupply_peers_g|id_district+x_quarter+x_young+x_parity,data=mydata,notes=FALSE)
       bias<-summary(res)$coeftable["x_lsupply_peers_g", "Estimate"]
      df_results$bias[i] <- bias-0.2
   # 2 recenter
      mydata$z=mydata$x_lsupply_peers_g-mydata$x_lsupply_peers_d
       res<-feols(y_lsupply~z|id_district+x_quarter+x_young+x_parity,data=mydata,notes=FALSE)
       bias<-summary(res)$coeftable["z", "Estimate"]
      df_results_recenter$bias[i] <- bias-0.2
    # Print
        if (i%%50==0){
               cat(paste(i,"\n"))
        }
        else{
               cat("*")
        }
}
df<-rbind(df_results,df_results_recenter)
datasummary_balance(bias~type,fmt = fmt_sprintf("%.3f"),dinm_statistic="p.value",df)
```

Show chart
```{r}
#| label: chart-bias
#| include: true
#| code-fold: true
#| warning: false
#| 



ggplot(df,aes(x=bias,group=type,fill=type))+
    geom_density(alpha=0.2)+
  scale_fill_viridis_d() +
  theme_minimal()+labs(color=" ",fill=" ")+
  theme(legend.position = "top")+ geom_vline(xintercept=0)
```




# 5 Predicing mother groups


## algorithm to predict new mother groups

```{r}
#| label: outcome-predict-fct
#| include: true
#| code-fold: true
#| warning: false

set.seed(1909)
# Define programme
predicted_peersupply<-function(dataframe,range=7){
  dt<-dataframe
  # Original ordering
  dt[, order := .I]
 # Randomly change date of birth 
  dt$x_dob_random<-dt$x_dob+sample(-range:range,nrow(dt),replace=TRUE)
 # Redo mother groups
  setorder(dt,id_district,  x_dob_random)
  dt[, count := 1:.N, by = id_district]
  dt[, group_id := floor((count - 1) / 5)]
  dt[, id_mother_group_random := paste(id_district, group_id, sep = "_")]
  # Leave-one-out peer effect within random mother group
      dt[, `:=`(
        sum_lsupply = sum(x_lsupply_base),
        n_group = .N
      ), by = id_mother_group_random]
      dt[, x_lsupply_peers_gr := (sum_lsupply - x_lsupply_base) / (n_group - 1)]
  # Recentered iv
      dt[,x_lsupply_peers_grc  := x_lsupply_peers_g-x_lsupply_peers_gr ]
  # Return
    # back to original order
    setorder(dt, order)
    return(dt$x_lsupply_peers_grc)
}
# testing

 dt<-DGP_fast()
 

dt$z<-predicted_peersupply(dt,3)

 feols(y_lsupply~z+x_age+x_income+x_schooling|id_district+x_quarter+x_young+x_parity,data=dt,notes=FALSE)

```


## simulate

```{r}
#| label: outcome-predict-sim
#| include: true
#| code-fold: true
#| warning: false
#| cache: true
  
# Iterations
I=100
results<-data.table(id=rep(1:nrow(dt),each=I),iteration=rep(1:I,each=nrow(dt)))
for (i in 1:I){

results[(((i - 1) * nrow(dt)) + 1): (i * nrow(dt)), y := predicted_peersupply(dt,14)]

}
collapsed <- results[, .(
  mean = mean(y, na.rm = TRUE),
  sd = sd(y, na.rm = TRUE)
), by = id]

 hist(collapsed$mean,breaks=50)
```


```{r}
#| label: outcome-sim_predicted
#| include: true
#| code-fold: true
#| warning: false
#| cache: true

#Iterations
II=1000
# Dataset
df_results<-data.frame(i=1:II,bias=NA,type="1 Conventional")
df_results_recenter<-data.frame(i=1:II,bias=NA,type="2 Recentered")


for (i in 1:II){
      # Generate DGP
    mydata<-DGP_fast()
     # Original ordering
     mydata[, order := .I]
    # 1 conventional reg
       res<-feols(y_lsupply~x_lsupply_peers_g|id_district+x_quarter+x_young+x_parity,data=mydata,notes=FALSE)
       bias<-summary(res)$coeftable["x_lsupply_peers_g", "Estimate"]
      df_results$bias[i] <- bias-0.2
      
      # 2 recenter 
        # Iterations
        I=10
        results<-data.table(order=rep(mydata$order,I),iteration=rep(1:I,each=nrow(dt)))
        for (j in 1:I){
        
        results[(((j- 1) * nrow(dt)) + 1): (j * nrow(dt)), y := predicted_peersupply(mydata,14)]
        
        }
        collapsed <- results[, .(
          mean = mean(y, na.rm = TRUE)
        ), by = order]
       mydata$x_lsupply_peers_rc<-collapsed$mean
    


      res<-feols(y_lsupply~x_lsupply_peers_rc|id_district+x_quarter+x_young+x_parity,data=mydata,notes=FALSE)
       bias<-summary(res)$coeftable["x_lsupply_peers_rc", "Estimate"]
      df_results_recenter$bias[i] <- bias-0.2


    # Print
        if (i%%50==0){
               cat(paste(i,"\n"))
        }
        else{
               cat("*")
        }
}
df<-rbind(df_results,df_results_recenter)
datasummary_balance(bias~type,fmt = fmt_sprintf("%.3f"),dinm_statistic="p.value",df)
```


Show chart
```{r}
#| label: chart-bias-predicted
#| include: true
#| code-fold: true
#| warning: false
#| 



ggplot(df,aes(x=bias,group=type,fill=type))+
    geom_density(alpha=0.2)+
  scale_fill_viridis_d() +
  theme_minimal()+labs(color=" ",fill=" ")+
  theme(legend.position = "top")+ geom_vline(xintercept=0)
```