---
title: "Simulating the mother groups DGP "
date: "2025-03-27"
author: "Hans Henrik Sievertsen"
format:
  html:
    code-fold: true
    code-summary: "Show the code"
---


# 1 Goal

The goal is to simulate a data generating process (DGP) mimicking the mother group assignment in Denmark. It is used to understand the role of fixed effects and the role of exclusion bias.

# 2 Preparation

We set seed, clear the workspace and load libraries. 

##### Set seed
```{r}
#| label: set-seed
#| include: true
#| code-fold: false

rm(list=ls())
set.seed(1909)
```

##### Load libriaries
```{r}
#| label: load-libraries
#| include: true
#| code-fold: false
#| warning: false
#| message: false

library(ggplot2)      # To create charts
library(dplyr)        # Data manipulation tools
library(lubridate)    # Date manipulation tools
library(fixest)       # Estimation
library(data.table)   # Data manipulation tools
library(modelsummary) # Create tables 
library(lmtest)       # Hypotheses testing
```




# 2 Define Data Generating Process

We here create a function that simulates a dataset mimicking the DGP we have in mind. The key features we focus on are

1. The role of unobservables (in group assignments and outcomes)

2. The assignment mechanism

3. The peer effects

**Re 1.**: We assume there is an individual unobservable $U$ that is a combination of  a unique individual effect and a district level effect. This unobservable affects variables like labor supply, income, parental labor supply and age. But *not* parity and child dob. 

**Re 2.**: The assignment is only strict based on district. THat is, *within each district*, nurses *sort* the mothers by

1. Child date of birth (days)

2. Mother age (integers)

3. Parity (binary)

We add a bit of noise to the sorting to mimick the real world where there will be errors, mothers coming in during the process etc. The nurse then simply starts filling mother groups by counting to five. That is, we *restrict mother groups to the size of 5*.

**Re 3.**: True peer effects are tricky to simulate. Here we cheat a bit by ignoring the multiplication effect/reflect effect. That is we:

- Define a latent level of mother labor supply depending on grand parent labor supply, the unobservables, etc. 

- Calculate the mother group leave one out average latent labor supply. 

- Calculate the *actual* labour supply by adding the peer effect to the latent effect.


That is it for now. 


```{r}
#| label: define-DGP
#| include: true
#| code-fold: true
#| warning: false

DGP_fast <- function(n = 70000, n_districts = 500) {
    
    # District identifier
    id_district <- sample(1:n_districts, n, replace = TRUE)
    # District unobserved effect
    district_effect <- rnorm(n_districts, mean = 0, sd = 1)
    # Individual unobserved effect
    U <- district_effect[id_district] + rnorm(n)
    # Covariates, note that parity and dob are unrelated to u, age, education, income, gm labour supply are not 
    x_age <- round(18 + 10 * runif(n) + 2 * U + rnorm(n))  # now depends on U
    x_age <- pmax(pmin(x_age, 40), 18)  # keep age between 18 and 40
    x_parity <- sample(1:2, n, replace = TRUE)
    x_schooling <- round(9 + 0.2 * x_age + 0.8 * U + rnorm(n))
    x_income <- round(200000 + 5000 * x_schooling + 10000 * U + rnorm(n, 0, 20000))
    x_gm_lsupply <- pmax(0, 20 + 0.5 * U + rnorm(n, 0, 5))
    x_dob <- sample(seq(as.Date("2010-01-01"), as.Date("2017-12-31"), by = "day"), n, replace = TRUE)
    # Create data frame
    dt <- data.table(id_mother = 1:n, id_district, x_age, x_parity,
                     x_schooling, x_income, x_gm_lsupply, x_dob)
    # Assign mother groups
    setorder(dt, id_district, x_dob, x_age,x_parity)
    dt[, order_original := .I , by = id_district]
    # Add small random noise to row numbers
    dt[, order_noisy := .I + rnorm(.N, mean = 0, sd = 1)]
    # Sort by the noisy key
    setorder(dt, id_district,order_noisy)
    dt[, order_noisy := .I , by = id_district]
    # Assign groups
    dt[, count := 1:.N, by = id_district]
    dt[, group_id := floor((count - 1) / 5)]
    dt[, id_mother_group := paste(id_district, group_id, sep = "_")]
    
    # Create fixed effects
    dt[, x_quarter := lubridate::quarter(x_dob, with_year = TRUE)]
    dt[, x_month := lubridate::month(x_dob)]
    dt[, x_year := lubridate::year(x_dob)]
    dt[, x_young := as.integer(x_age < 30)]
    dt[, x_age5 := floor(x_age/5)*5]
    # Leave-one-out variables
    dt[, `:=`(
      sum_schooling = sum(x_schooling),
      sum_income = sum(x_income),
      sum_gm=sum(x_gm_lsupply),
      n_group = .N
    ), by = id_mother_group]
    dt[, x_schooling_g := (sum_schooling - x_schooling) / (n_group - 1)]
    dt[, x_income_g := (sum_income - x_income) / (n_group - 1)]
    dt[, x_gm_lsupply_g := (sum_gm - x_gm_lsupply) / (n_group - 1)]
    
    # Create labor supply (our main outcome)
    # Initial (baseline) labor supply without peer effect
    dt[, lsupply_base := pmin(40, pmax(0, 
                                       20 +         # base level
                                         0.5 * U +     # unobserved factor
                                         0.3 * x_gm_lsupply +  # grandmother's influence
                                         rnorm(.N, 0, 5)  # noise
    ))]
    # Leave-one-out peer effect within mother group
    dt[, `:=`(
      sum_lsupply = sum(lsupply_base),
      n_group = .N
    ), by = id_mother_group]
    
    dt[, x_lsupply_g := (sum_lsupply - lsupply_base) / (n_group - 1)]
    # Final labor supply including peer effect
    dt[, y_lsupply := pmin(40, pmax(0,
                                    lsupply_base +
                                      0.2 * x_lsupply_g  # peer effect coefficient
    ))]
    
    # Keep only relevant columns
    dt<-dt[, .SD, .SDcols = grep("^id_|^x_|^gm_|^y_", names(dt), value = TRUE)]
      return(dt)
}

```


#### Show data
```{r}
#| label: show data
#| include: true
#| code-fold: true
#| warning: false

df<-DGP_fast()
setorder(df, id_district, x_parity, x_dob, x_age)
head(df[, .SD, .SDcols = c("id_district","id_mother_group","y_lsupply")],n=11)


```

# 3 Check that this looks like our real data

## 3.1 Balance checks showing the problem

### 3.1.1 Approach 1

```{r}
#| label: balance-one regressions
#| include: true
#| code-fold: true
#| warning: false

summary(feols(x_gm_lsupply~x_gm_lsupply_g|id_district+x_age5,data=df))

```


Okay this looks similar 

Let us try to randomize groups and recalculate and check distribution

```{r}
#| label: balance-one regressions
#| include: true
#| code-fold: true
#| warning: false

#Iterations
I=10000
# Dataset
df_sim<-data.frame(i=1:I,beta=NA,type="Raw")

for (i in 1:I){
  # random order
  df[, random :=rnorm(.N, mean = 0, sd = 1)]
  # reorder
  setorder(df, id_district, x_age5,random)
  # Assign groups
  df[, count := 1:.N, by = id_district]
  df[, group_id := floor((count - 1) / 5)]
  df[, id_mother_group_new := paste(id_district, group_id, sep = "_")]
  # Recalculate leave one out mean
   df[, `:=`(
      sum_gm=sum(x_gm_lsupply),
      n_group = .N
    ), by = id_mother_group_new]
    df[, x_gm_lsupply_gshuffle := (sum_gm - x_gm_lsupply) / (n_group - 1)]
    
  # Save estimate
    res<-feols(x_gm_lsupply~x_gm_lsupply_gshuffle|id_district+x_age5,data=df,notes=FALSE)
    # Extract coefficient
    beta <- summary(res)$coeftable["x_gm_lsupply_gshuffle", "Estimate"]
    # Store beta-value in results dataset
    df_sim$beta[i] <- beta
    
      # Print
        if (i%%50==0){
               cat(paste(i,"\n"))
        }
        else{
               cat("*")
        }
    
}

```


Add actual to chart


```{r}
#| label: balance-plot
#| include: true
#| code-fold: true
#| warning: false


# Store main
main<-feols(x_gm_lsupply~x_gm_lsupply_g|id_district+x_age5,data=df)
beta <- summary(main)$coeftable["x_gm_lsupply_g", "Estimate"]



ggplot(df_sim,aes(x=beta))+geom_histogram(color="black",fill="white") +
  theme_minimal()+geom_vline(xintercept = beta)
```

### 3.1.2 Approach 2


```{r}
#| label: reg on cells


# datafarme to save results
df_results<-data.table(id_district=rep(unique(df$id_district),times=4),
                       p=NA,
                       var=rep(c("x_schooling","x_income","x_age", "x_gm_lsupply", "x_parity"),each=length(rep(unique(df$id_district))))
                               )
# Loop over variables
for (v in c("x_schooling","x_income","x_age", "x_gm_lsupply", "x_parity")){
  # Loop over all districts:
  for (i in unique(df$id_district)){
    # select sample
    filtered_df <- df[df$id_district== i, ]
     
    # Model without fixed effects
    mod1 <- feols(as.formula(paste(v, "~ 1")), data = filtered_df)
    
    # Model with fixed effects
    mod2 <- feols(as.formula(paste(v, "~ 1 | id_mother_group_new")), data = filtered_df)
    
    # Compare models
    ll1 <- logLik(mod1)
    ll2 <- logLik(mod2)
    lr_stat <- 2 * (as.numeric(ll2) - as.numeric(ll1))
    df_fe <- length(unique(filtered_df$id_mother_group_new)) - 1
    pval <- pchisq(lr_stat, df = df_fe, lower.tail = FALSE)
    # Store results
    df_results$p[df_results$id_district == i & df_results$var == v] <- pval
  }
  print(v)
}

```

Now create plot

```{r}

# Sort by var and p
setorder(df_results, var, p)

# Compute relative rank within each var group
df_results[, relrank := rank(p, ties.method = "first") / .N, by = var]


# Plot
library(ggplot2)
ggplot(df_results, aes(x = relrank, y = p, color = var)) +
  geom_line() +
  labs(x = "Relative Rank", y = "p-value", color = "Variable") +
  theme_minimal()
```

Now within age and district cells


```{r}
df[, group := .GRP, by = .(id_district,x_age5)]

# datafarme to save results
df_results<-data.table(group=rep(unique(df$group),times=5),
                       p=NA,
                       var=rep(c("x_schooling","x_income","x_age", "x_gm_lsupply", "x_parity"),each=length(rep(unique(df$group))))
                               )
# Loop over variables
for (v in c("x_schooling","x_income","x_age", "x_gm_lsupply")){
  # Loop over all districts:
  for (i in unique(df$group)){
    # select sample
    filtered_df <- df[df$group== i, ]
    
    # Skip if the variable is constant or NA
    y <- filtered_df[[v]]
    if (length(unique(y[!is.na(y)])) <= 3) next  # Skip constant or all-NA
    
    
    # Model without fixed effects
    mod1 <- feols(as.formula(paste(v, "~ 1")), data = filtered_df)
    
    # Model with fixed effects
    mod2 <- feols(as.formula(paste(v, "~ 1 | id_mother_group_new")), data = filtered_df)
    
    # Compare models
    ll1 <- logLik(mod1)
    ll2 <- logLik(mod2)
    lr_stat <- 2 * (as.numeric(ll2) - as.numeric(ll1))
    df_fe <- length(unique(filtered_df$id_mother_group_new)) - 1
    pval <- pchisq(lr_stat, df = df_fe, lower.tail = FALSE)
    # Store results
    df_results$p[df_results$group == i & df_results$var == v] <- pval
  }
  print(v)
}

```

Now create plot

```{r}
# Remove rows with NA in p
df_results <- df_results[!is.na(p)]

# Sort by var and p
setorder(df_results, var, p)

# Compute relative rank within each var group
df_results[, relrank := rank(p, ties.method = "first") / .N, by = var]


# Plot
library(ggplot2)
ggplot(df_results, aes(x = relrank, y = p, color = var)) +
   geom_segment(aes(x = 0, xend = 1, y = 0, yend =1),color="black")+
  geom_line() +
  labs(x = "Relative Rank", y = "p-value", color = "Variable") +
  theme_minimal()+
  xlim(0,1)+ylim(0,1)
```



## 3.2 Balance checks showing the solution A: control for pool level


```{r}
#| label: balance-one-sol
#| include: true
#| code-fold: true
#| warning: false

#Iterations
I=10000
# Dataset
df_sim_pool<-data.frame(i=1:I,beta=NA,type="Raw")

 # Calculate pool level leave one out mean
  df[, group := .GRP, by = .(id_district)]
   df[, `:=`(
      sum_gm=sum(x_gm_lsupply),
      n_group = .N
    ), by = group]
  df[, x_gm_lsupply_da := (sum_gm - x_gm_lsupply) / (n_group - 1)]
    
  # Check regs
  
  feols(x_gm_lsupply~x_gm_lsupply_g|id_district+x_age5,data=df,notes=FALSE)
  feols(x_gm_lsupply~x_gm_lsupply_g+x_gm_lsupply_da|id_district+x_age5,data=df,notes=FALSE)
  
  
for (i in 1:I){
  # random order
  df[, random :=rnorm(.N, mean = 0, sd = 1)]
  # reorder
  setorder(df, id_district, x_age5,random)
  # Assign groups
  df[, count := 1:.N, by = id_district]
  df[, group_id := floor((count - 1) / 5)]
  df[, id_mother_group_new := paste(id_district, group_id, sep = "_")]
  # Recalculate leave one out mean
   df[, `:=`(
      sum_gm=sum(x_gm_lsupply),
      n_group = .N
    ), by = id_mother_group_new]
    df[, x_gm_lsupply_gshuffle := (sum_gm - x_gm_lsupply) / (n_group - 1)]
    
  # Save estimate
    res<-feols(x_gm_lsupply~x_gm_lsupply_gshuffle+x_gm_lsupply_da|id_district+x_age5,data=df,notes=FALSE)
    # Extract coefficient
    beta <- summary(res)$coeftable["x_gm_lsupply_gshuffle", "Estimate"]
    # Store beta-value in results dataset
    df_sim_pool$beta[i] <- beta
    
      # Print
        if (i%%50==0){
               cat(paste(i,"\n"))
        }
        else{
               cat("*")
        }
    
}

```


Add actual to chart


```{r}
#| label: balance-plot
#| include: true
#| code-fold: true
#| warning: false


# Store main
main<-feols(x_gm_lsupply~x_gm_lsupply_g+x_gm_lsupply_da|id_district+x_age5,data=df)
beta <- summary(main)$coeftable["x_gm_lsupply_g", "Estimate"]



ggplot(df_sim_pool,aes(x=beta))+geom_histogram(color="black",fill="white") +
  theme_minimal()+geom_vline(xintercept = beta)
```




## 3.2 Balance checks showing the solution B: recenter

```{r}
#| label: recenter
#| include: true
#| code-fold: true
#| warning: false


df_test<-df

 setorder(df_test, id_mother)

 I=1000
# Empty data table to store results
df_permute<-data.table(id_mother=rep(df_test$id_mother,times=I),
                        iteration=rep(1:I,each=length(df_test$id_mother),
                        beta=NA)
                  )
                       
                       
# simulate
for (i in 1:I){
  # New data of birth
    df_test[, x_dob_new := x_dob + round(runif(.N, min = -14, max = 14))]
  # Assign mother groups
    setorder(df_test, id_district, x_dob_new, x_age,x_parity)
    df_test[, order_original := .I , by = id_district]
  # Assign groups
    df_test[, count := 1:.N, by = id_district]
    df_test[, group_id := floor((count - 1) / 5)]
    df_test[, id_mother_group_new := paste(id_district, group_id, sep = "_")]
   # Recalculate leave one out mean
   df_test[, `:=`(
      sum_gm=sum(x_gm_lsupply),
      n_group = .N
    ), by = id_mother_group_new]
    df_test[, x_gm_lsupply_gs := (sum_gm - x_gm_lsupply) / (n_group - 1)]
   # Save
    setorder(df, id_mother)
    df_permute$beta[df_permute$iteration == i ] <- df_test$x_gm_lsupply_gs
    
          # Print
        if (i%%50==0){
               cat(paste(i,"\n"))
        }
        else{
               cat("*")
        }
    
}
df_collapsed <- df_permute[, .(x_gm_lsupply_gsim = mean(beta)), by = id_mother]


# now merge on
  df<-merge(df,df_collapsed,by="id_mother")
  df[,x_gm_lsupply_grc :=x_gm_lsupply_g-x_gm_lsupply_gsim ]
  
I=10000
# Dataset
df_sim_recent<-data.frame(i=1:I,beta=NA,type="Raw")


for (i in 1:I){
  # random order
  df[, random :=rnorm(.N, mean = 0, sd = 1)]
  # reorder
  setorder(df, id_district, x_age5,random)
  # Assign groups
  df[, count := 1:.N, by = id_district]
  df[, group_id := floor((count - 1) / 5)]
  df[, id_mother_group_new := paste(id_district, group_id, sep = "_")]
  # Recalculate leave one out mean
   df[, `:=`(
      sum_gm=sum(x_gm_lsupply),
      n_group = .N
    ), by = id_mother_group_new]
    df[, x_gm_lsupply_gshuffle := (sum_gm - x_gm_lsupply) / (n_group - 1)]
    
    
      df[,x_gm_lsupply_gshufflerc :=x_gm_lsupply_gshuffle-x_gm_lsupply_gsim ]
  # Save estimate
    res<-feols(x_gm_lsupply~x_gm_lsupply_gshufflerc|id_district+x_age5,data=df,notes=FALSE)
    # Extract coefficient
    beta <- summary(res)$coeftable["x_gm_lsupply_gshufflerc", "Estimate"]
    # Store beta-value in results dataset
    df_sim_recent$beta[i] <- beta
    
      # Print
        if (i%%50==0){
               cat(paste(i,"\n"))
        }
        else{
               cat("*")
        }
    
}


```


```{r}
#| label: balance-plot-recenter
#| include: true
#| code-fold: true
#| warning: false


# Store main
main<-feols(x_gm_lsupply~x_gm_lsupply_grc|id_district+x_age5,data=df)
beta <- summary(main)$coeftable["x_gm_lsupply_grc", "Estimate"]



ggplot(df_sim_recent,aes(x=beta))+geom_histogram(color="black",fill="white") +
  theme_minimal()+geom_vline(xintercept = beta)
```
